Fractional Types

### 2019-10-30 fracGC/p.tex
a bit on intro

>>>>

\paragraph*{Conservation of Information.} If quantum field theory is
correct (as it so far seems to be) then information, during any
physical process, is neither created nor
destroyed. Landauer~\cite{Landauer:1961,Landauer,bennett1985fundamental},
Bennett~\cite{bennett2010notes,bennett2003notes,Bennett:1973:LRC},
Fredkin~\cite{fredkin1982conservative} and others made compelling
arguments that this physical principle induces a corresponding
computational principle of ``conservation of information.'' This
principle characterizes quantum computing and its classical
restriction known as reversible computing.

\paragraph*{Quantum (Reversible) Computing.} The quantum circuit model
can be characterized as a classical reversible circuit model augmented
with parallel operations modeling superpositions. Computations in the
classical submodel can be formalized as isomorphisms on finite types,
i.e., the types generated from the empty type 0, the unit type 1, and
sums and products. Put differently, the principle of conservation of
information is expressed as computating via sound and complete type
isomorphisms~\cite{James:2012:IE:2103656.2103667} or type
equivalences~\cite{DBLP:conf/esop/CaretteS16}.

\paragraph*{Negative Entropy.}  
According to the conventional theory of
communication~\cite{Shannon1948}, a type with $N$ values is viewed as
an abstract system that has $N$ distinguishable states where the
amount of information contained in each state is $\log{N}$. This
entropy is a measure of information which materializes itself in
memory or bandwidth requirements when storing or transmitting elements
of this type. Thus a type with 8 elements needs 3 bits of memory for
storage or 3 bits of bandwidth for communication. The logarithmic map
implies that information contained in a composite state is the sum of
the information contained in its constituents. For example, the type
$A \times B$ where $A$ has two elements and $B$ has three elements can
be thought of a composite system consisting of two independent
unrelated subsystems.  Each state of the composite system therefore
contains $\log{(2*3)} = \log{2} + \log{3}$ bits which is the sum of
the information contained in each subsystem. A \emph{fractional type}
$\frac{1}{A}$ introduces negative entropy. For example, a type with
cardinality $\frac{1}{8}$ has entropy $\log{\frac{1}{8}} = -3$. It is
natural to interpret this negative entropy just like we interpret
``negative money,'' as a resource (space or bandwidth) to be repaid
(reclaimed) by some other part of the system.

\paragraph*{Temporary Storage using Ancilla Bits.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivating Example}
